# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
*Load the data
```{r}
activity <- read.csv("activity.csv")
activity$interval <- as.factor(activity$interval)
sumup <- tapply(activity$steps,activity$date,sum)
sumup <- as.data.frame(sumup)
sumup$date <- as.Date(unique(activity$date))
```


## What is mean total number of steps taken per day?

*Make Plot
```{r}
library(lattice)
xyplot(sumup$sumup~sumup$date,type="h",xlab="Date",ylab="Steps Per Day")
```

*Calculate mean and median on daily basis
```{r}
mean <- mean(sumup$sumup,na.rm=TRUE)
median <- median(sumup$sumup,na.rm=TRUE)
```
The mean of total number of steps taken per day is `r mean`.
The median of total number of steps taken per day is `r median`.

## What is the average daily activity pattern?

*Calculate mean across all days & plot it
```{r}
s <- tapply(activity$steps,activity$interval,function(x) mean(x,na.rm=TRUE))
xyplot(s~activity$interval,type="l",xlab="Time Interval(5mins per tick)",ylab="averaged steps taken")
max <- max(s)
```

Maximum number of steps taken in one day: `r max`

## Imputing missing values
*Calculate total number of missing values
```{r}
missval <- sum(is.na(activity$steps))
```
The total number of missing values is `r missval`

*Replace all the missing values with average steps taken by interval
```{r}
for(i in 1:nrow(activity)){
        if(is.na(activity$steps[i])){
                activity$steps[i] <- s[which(rownames(s)==activity$interval[i])]
        }
}
```
Here's brief intro to this for loop: it tells if we find a missing value,  
look up its time interval and find the corresponding averaged steps taken,  
and then assign it the value.  

*Make Plot after imputing missing values
```{r}
sumup <- tapply(activity$steps,activity$date,sum)
sumup <- as.data.frame(sumup)
sumup$date <- as.Date(unique(activity$date))
xyplot(sumup$sumup~sumup$date,type="h",xlab="Date",ylab="Steps Per Day",)
```

*Calculate mean and median on daily basis after imputing missing values
```{r}
mean <- mean(sumup$sumup,na.rm=TRUE)
median <- median(sumup$sumup,na.rm=TRUE)
```
The new mean of total number of steps taken per day is `r mean`.
The new median of total number of steps taken per day is `r median`.

Apparently these mean and median are not different from what we've calculated before.

## Are there differences in activity patterns between weekdays and weekends?
*Create new columns indicating weekdays or weekends
```{r}
activity$date <- as.Date(activity$date)
activity$type <- NA
for(i in 1:nrow(activity)){
        if(weekdays(activity$date[i])=="周六" || weekdays(activity$date[i])=="周日"){
                activity$type[i] <- "weekend"
        }else {activity$type[i] <- "weekday"}
}
```

*Create plot
```{r}
s <- tapply(activity$steps,activity$interval,mean)
xyplot(s~activity$interval | activity$type,type="l",xlab="Time Interval(5mins per tick)",ylab="averaged steps taken",layout=c(1,2))
```



